---
title: "Data Tidying"
output: github_document
---
# load package
```{r}
library(tidyverse) # tidyverse is a range of packages


options(tibble.print_min = 5)

```


# pivot_longer
In data import, we used the haven package to load the PULSE biomarkers dataset from a .sas7bdat. Let’s reload those data and take a closer look:
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_df
```

With our new understanding of tidy data, we quickly recognize a problem: the BDI score is spread across four columns, which correspond to four observation times. We can fix this problem using pivot_longer:

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, #dataset
    bdi_score_bl:bdi_score_12m, #column range
    names_to = "visit", #new column name
    values_to = "bdi") # value

pulse_tidy_data
```

去掉“vist”值的前缀
This looks much better! However, now visit is an issue. The original column names were informative but we probably don’t need to keep the bdi_score_ prefix in each case. I’ll use an additional option in pivot_longer to address this:

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_", #remove prefix
    values_to = "bdi")

pulse_tidy_data
```

## using $>$ 
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% #load dataset
  janitor::clean_names() %>%  # clean dataset
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>% #For tidying single tables
  relocate(visit) %>% #list "visit" to be the first column
  mutate(
    visit = replace(visit, visit == "bl", "00m"), # replace the value "bl" by "00m" in "visit" column
    visit = factor(visit)) %>% # turns "visit" into factor
  arrange(id, visit) #arrange by "id" and "visit"

print(pulse_df, n = 12) #print the dataset with 12 rows
```

```{r, show_col_types = FALSE}

litters_wide = 
  read_csv("./data/FAS_litters.csv", show_col_types = FALSE) %>%
  janitor::clean_names()
litters_wide

litters_wide = 
  read_csv("./data/FAS_litters.csv", show_col_types = FALSE) %>%
  janitor::clean_names()%>%
  select(litter_number, ends_with("weight")) %>% 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd", 
    values_to = "weight") %>% 
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))
litters_wide
```


