---
title: "Data Manipulation"
output: github_document
---
# load package
```{r}
library(tidyverse) # tidyverse is a range of packages
```

# load in the FAS litters data
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

# 'select'

## choose/remove some columns

```{r}
select(litters_df, group, gd0_weight)

select(litters_df, - gd0_weight)

```

## selct a range column
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

## renaming columns
```{r}
# in one step
select(litters_df, GROUP = group, LITTER_NUMBER = litter_number)

# in multiple steps
rename(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

## select helpers
下面打?select_helpers

## select using keywords
```{r}
select(litters_df, starts_with("gd"))
select(litters_df, ends_with("ht"))
select(litters_df, contains("ht"))
```

## sequencing
把litter_number放在最前面但是保留其他的columns
everything(), which is handy for reorganizing columns without discarding anything
```{r}
select(litters_df, litter_number, everything())

```

OR
```{r}
relocate(litters_df, litter_number)

```

# 'filter'
You will often filter using comparison operators (>, >=, <, <=, ==, and !=). You may also use %in% to detect if values appear in a set, and is.na() to find missing values. The results of comparisons are logical – the statement is TRUE or FALSE depending on the values you compare – and can be combined with other comparisons using the logical operators & and |, or negated using !.

与select不同点：
1. select 处理column; filter处理rows
2. fiilter 有conditioners, based on ture/false, filter rows

## remove/keep rows

```{r}
filter(litters_df, gd_of_birth < 22)
```

```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, !(gd_of_birth == 20)) # flip the result
```

```{r}
filter(litters_df,  group %in% c("Con7", "Con8")) # keep groups = con7 or con8 

```
```{r}
filter(litters_df,  group == "Con7" & gd_of_birth == 20)
```

## filter and NA
Remove All NAS in the whole dataset
```{r}
drop_na(litters_df)
```
Remove NAS only in column "gd0_weight"
```{r}
drop_na(litters_df, gd0_weight) 
```
# 'mutate'

modify old variables or create new variables
```{r}
mutate(litters_df,   wt_gain = gd18_weight - gd0_weight, #muate a new variable#
  group = str_to_lower(group)) #modify the "group" variable with all lower cases
```

# 'arrange'
put things in orders

```{r}
head(arrange(litters_df, group, pups_born_alive), 10) # show first 10 rows
# arrange the litters_df by group and pups_born_alive
```

# '%>%' data processing

## bad example
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```
```{r}
litters_data_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )
```

## Using %>% 
%>% = "then"
%>% 快捷键：Cmd + Shift + M 

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
```

```{r}
read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>% 
  filter(sex == 1) %>% 
  select(-pd_ears) %>% 
  mutate(pd_pivot_gt7 = pd_pivot > 7)
```

## Using . as placeholder
You can make this more explicit by using . as a placeholder for the result of the preceding call
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(data = ., wt_gain)
```

